Equation-free analysis of a dynamically evolving multigraph

### 2016-02-07 report/multigraph_manuscript.tex
Things continue to work, must document. Working on manuscript text.

\title{Equation-free analysis of a dynamically evolving multigraph}
\author{Alexander Holiday \inst{1} \and Bal\'{a}zs R\'{a}th \inst{2}
  \and Ioannis G. Kevrekidis \inst{1.3}} \institute{Department of
  Chemical and Biological Engineering, Princeton University,
  Princeton, New Jersey 08544, USA \and Institute of Mathematics,
  Budapest University of Technology, 1 Egry J\'{o}zsef u., Budapest
  1111, Hungary \and PACM, Princeton University, Princeton, New Jersey
  08544, USA}
% $^1$\textit{Department of Chemical and Biological Engineering},
% \textit{Princeton University, Princeton, New Jersey 08544, USA} \\ \\
% $^2$\textit{Department of Mathematics}, \\
% \textit{The University of British Columbia, Vancouver, British
%   Columbia V6T 1Z4} } \date{}

@@

\begin{figure}[th!]
  \vspace{-5mm}
  \centering
  \begin{subfigure}[t]{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{triangle-slaving}
    \subcaption{\label{fig:sv-n2}}
  \end{subfigure} %
  \begin{subfigure}[t]{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{triangle-slaving-n3}
    \subcaption{\label{fig:sv-n3}}
  \end{subfigure}
  \caption{Evolution of higher-order statistics (here, the total
    triangle count) for three different network initializations. They
    are quickly ($O(n^2)$ steps) drawn to a slow manifold on which
    they slowly evolve ($O(n^3)$ steps) to a steady
    state. \label{fig:sv}}
\end{figure}

@@

The pieces described above are combined to accelerate simulations through coarse-projective integration. Denoting the lifting operator by $\mathbf{L} \; : \; c \rightarrow G$ where $c \in \mathbb{R}^7$ is the vector of polynomial coefficients, and the restriction operator as $\mathbf{R} \; : G \rightarrow c$, the method progresses as follows:

@@

A representative evolution of one of the polynomial coefficients and its projected value is shown in Fig. (\ref{fig:coeff-evo}). Note that Step (\ref{cpi:heal}) is necessary to allow the system to approach its slow manifold. Upon initializing a new network in Step (\ref{cpi:init}), certain variables will be far from their expected values in a normal system; however, by hypothesis these quantities quickly evolve to a state in which they are strongly correlated with, or are functions of, the selected coarse variables. Thus, after a short interval of ``healing'', they are drawn to the expected trajectory, after which we begin sampling. This is analogous to Fig. (\ref{fig:sv}). The computational gains arise from the projective step, (\ref{cpi:proj}). Here, we advance the system $t_p$ steps at the cost of one evaluation of forward Euler, instead of $t_p$ direct evaluations. \\

@@

\begin{figure}[h!]
  \vspace{-5mm}
  \centering
  \begin{subfigure}{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{cpi-3d-comp}
    \subcaption{\label{fig:cpi-error}}
  \end{subfigure} %
  \begin{subfigure}{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{cpi-relative-error-n3}
    \subcaption{\label{fig:self-error}}
  \end{subfigure}%
  \caption{CPI results: (a) the evolution of the CPI-accelerated
    degree sequence (red) compared to direct simulation (blue) and (b)
    error in cpi-accelerated runs calculated by comparing
    cpi-accelerated degree sequences to those arising from an ensemble
    of direct simulations. \label{fig:cpi-results}}
\end{figure}

@@

When applying machine learning techniques to a dataset, it is necessary to define a distance (or similarity) between each data point. If these points are a set of vectors in $\mathbb{R}^n$ one has a number of metrics to choose from, the Euclidean distance being a common first option. Unfortunately, when individual points are not vectors but networks, the definition of a computationally feasible metric becomes far more challenging. Examples such as the maximal common subgraph and edit distances, defined in \cite{bunke_graph_1998} and \cite{gao_survey_2010} do define metrics on the space of graphs, but their computation is $NP\textnormal{-}hard$. \\

The strategy used in the following methods, detailed in \cite{rajendran_analysis_2013}, is to embed each graph based on statistics of random walks along its edges and vertices. This maps each network to a vector in $\mathbb{R}^n$, and the Euclidean distance is subsequently used to calculate the similarity of two graphs. Full details are given in (\cite{rajendran_analysis_2013}).

@@

\begin{figure}[h!]
  \vspace{-5mm}
  \centering
  \begin{subfigure}{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{dmaps-3d-convergence}
    \subcaption{\label{fig:dmaps-results-regular}}
  \end{subfigure} %
  \begin{subfigure}{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{dmaps-3d-convergence-zoom}
    \subcaption{\label{fig:dmaps-results-zoom}}
  \end{subfigure}%
  \caption{DMAP of motif-based embeddings from two separate simulation
    trajectories: (a) embedding of two trajectories on different
    timescales. While the trajectories are initially embedded in
    separate regions due to their different initial conditions,
    they're eventually mapped close together as they evolve to the
    same coarse stationary state. Final states are circled, and point
    size grows as more steps are taken in each trajectory. (b)
    Enlarged view of the final points in the previous plot, revealing
    the approach to a similar final state. Here, the average of the
    final fifty points is circled. Due to the stochastic nature of
    the system, the final embeddings randomly fluctuate about the
    shared stationary state. \label{fig:dmaps-results}}
\end{figure}

@@

>>>>

\title{Equation-free analysis of a dynamically evolving multigraph}
\author{Alexander Holiday \inst{1} \and Ioannis G. Kevrekidis
  \inst{1,2}} \institute{Department of Chemical and Biological
  Engineering, Princeton University, Princeton, New Jersey 08544, USA
  \and PACM, Princeton University, Princeton, New Jersey 08544, USA}
% $^1$\textit{Department of Chemical and Biological Engineering},
% \textit{Princeton University, Princeton, New Jersey 08544, USA} \\ \\
% $^2$\textit{Department of Mathematics}, \\
% \textit{The University of British Columbia, Vancouver, British
%   Columbia V6T 1Z4} } \date{}

@@

\begin{figure}[th!]
  % \vspace{-5mm}
  % \centering
  % \begin{subfigure}[t]{0.49\textwidth}
  %   \centering
  %   \includegraphics[width=\textwidth]{triangle-slaving}
  %   \subcaption{\label{fig:sv-n2}}
  % \end{subfigure} %
  % \begin{subfigure}[t]{0.49\textwidth}
    \centering
    \includegraphics[width=0.8\textwidth]{triangle-slaving-n2-n3}
    % \subcaption{\label{fig:sv-n3}}
  % \end{subfigure}
  \caption{Evolution of higher-order statistics (here, the total
    triangle count) for three different network initializations. They
    are quickly ($O(n^2)$ steps) drawn to a slow manifold on which
    they slowly evolve ($O(n^3)$ steps) to a steady
    state. \label{fig:sv}}
\end{figure}

@@

The pieces described above are combined to accelerate simulations through coarse-projective integration. Denoting the lifting operator by $\mathbf{L} \; : \; c \rightarrow G$ where $c \in \mathbb{R}^4$ is the vector of polynomial coefficients, and the restriction operator as $\mathbf{R} \; : G \rightarrow c$, the method progresses as follows:

@@

Note that Step (\ref{cpi:heal}) is necessary to allow the system to approach its slow manifold. Upon initializing a new network in Step (\ref{cpi:init}), certain variables will be far from their expected values in a normal system; however, by hypothesis these quantities quickly evolve to a state in which they are strongly correlated with, or are functions of, the selected coarse variables. Thus, after a short interval of ``healing'', they are drawn to the expected trajectory, after which we begin sampling. This is analogous to Fig. (\ref{fig:sv}). The computational gains arise from the projective step, (\ref{cpi:proj}). Here, we advance the system $t_p$ steps at the cost of one evaluation of forward Euler, instead of $t_p$ direct evaluations. \\

@@

\begin{figure}[h!]
  \vspace{-5mm}
  \centering
  \begin{subfigure}{0.59\textwidth}
    \centering
    \includegraphics[width=\textwidth]{cpi-3d-comp-stretch}
    \subcaption{\label{fig:cpi-error}}
  \end{subfigure} %
  \begin{subfigure}{0.39\textwidth}
    \centering
    \includegraphics[width=\textwidth]{cpi-relative-error-n3}
    \subcaption{\label{fig:self-error}}
  \end{subfigure}%
  \caption{CPI results: (a) the evolution of the CPI-accelerated
    degree sequence (red) compared to direct simulation (blue) and (b)
    error in cpi-accelerated runs calculated by comparing
    cpi-accelerated degree sequences to those arising from an ensemble
    of direct simulations. \label{fig:cpi-results}}
\end{figure}

@@

When applying machine learning techniques to a dataset, it is
necessary to define a distance (or similarity) between each pair of
data point. If these points are a set of vectors in $\mathbb{R}^n$ one
has a number of metrics to choose from, the Euclidean distance being a
common first option. Unfortunately, when individual points are not
vectors but networks, the definition of a computationally feasible
metric becomes far more challenging. Examples such as the maximal
common subgraph and edit distances, defined in \cite{bunke_graph_1998}
and \cite{gao_survey_2010} do define metrics on the space of graphs,
but their computation is $NP\textnormal{-}hard$. Other computationally
feasible approaches include comparing distributions of random
walks on each graph \cite{kondor}, calculating the so-called
$n$-tangle density \cite{fefferman}, or calculting the edit distance
with an approximation algorithm \cite{riesen,zeng}.  \\

The strategy used in the following methods, detailed in
\cite{rajendran_analysis_2013} and \cite{gu} enumerates the number of
times a certain set of motifs (or subgraphs) appears in each network
in the dataset. This maps each network to a vector in $\mathbb{R}^n$,
and the Euclidean distance is subsequently used to calculate the
similarity of two graphs. Due to computational restrictions, we chose
to count the number of three- and four-vertex subgraphs that were
contained in each network. As there are eight such motifs, shown in
Fig. (\ref{fig:motifs}), this process $\gamma$ maps each graph to an
eight-dimensional vector: $\gamma : G \rightarrow
\mathbb{R}^8$. Additionally, we applied this operateration to a
simplified version of each graph, wherein any multiple edges were
reduced to single edges.

\begin{figure}[h!]
  \vspace{-5mm}
  \centering
  \includegraphics[width=\textwidth]{motifs}
  \caption{List of the subgraphs used to embed each network. The
    number of times each appeared in the simplified input graph was
    calculated, effectively mapping graphs to $\mathbb{R}^8$. \label{fig:motifs}}
\end{figure}

@@

\begin{figure}[h!]
  \vspace{-5mm}
  \centering
  \begin{subfigure}{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{dmaps-3d-convergence-a}
    \subcaption{\label{fig:dmaps-results-regular}}
  \end{subfigure} %
  \begin{subfigure}{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{dmaps-3d-convergence-zoom-a}
    \subcaption{\label{fig:dmaps-results-zoom}}
  \end{subfigure}%
  \caption{DMAP of motif-based embeddings from two separate simulation
    trajectories: (a) embedding of two trajectories on different
    timescales. While the trajectories are initially embedded in
    separate regions due to their different initial conditions,
    they're eventually mapped close together as they evolve to the
    same coarse stationary state. Final states are circled, and point
    size grows as more steps are taken in each trajectory. (b)
    Enlarged view of the final points in the previous plot, revealing
    the approach to a similar final state. Here, the average of the
    final fifty points is circled. Due to the stochastic nature of
    the system, the final embeddings randomly fluctuate about the
    shared stationary state. \label{fig:dmaps-results}}
\end{figure}

@@

\begin{acknowledgement}
The authors thank Bal\'{a}zs R\'{a}th for his helpful discussion and
insightful suggestions that have greatly enhanced this work.
\end{acknowledgement}
